Model-based variable selection for regression with highly correlated variables.

### 2014-08-22 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@814 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

However, $\mathcal{S}$ being combinatorial, $|\mathcal{S}|$ is huge. It has two cumulated consequences: First, the {\it exact} probability $\mathbb{P}(\boldsymbol{S}|\boldsymbol{X})$ may be of the same order of magnitude for a large number of candidates  $\boldsymbol{S}$, including the best one; Second, the {\sc bic} {\it approximations} of this quantity may introduce additional confusion to wisely distinguish between model probabilities. In order to limit this problem, we propose to introduce some information in $\mathbb{P}(\boldsymbol{S})$ promoting simple models through the following {\it hierarchical} uniform distribution denoted by $\mathbb{P}_H(\boldsymbol{S})$:
\begin{eqnarray}
\mathbb{P}_H(\boldsymbol{S}) & = & \mathbb{P}_H(\boldsymbol{J}_r,\boldsymbol{J}_p) \\
 & = & \mathbb{P}_H(\boldsymbol{J}_r,\boldsymbol{J}_p,d_r,\boldsymbol{d}_p) \\
 & = & \mathbb{P}_U(\boldsymbol{J}_p|\boldsymbol{d}_p,\boldsymbol{J}_r,d_r) \times \mathbb{P}_U(\boldsymbol{d}_p|\boldsymbol{J}_r,d_r) \times \mathbb{P}_U(\boldsymbol{J}_r|d_r)\times \mathbb{P}_U(d_r) \\
 & = & \left[\prod_{j=1}^{d_r} \left(\begin{array}{c} d-d_r \\ d_p^j \end{array}\right) \right]^{-1} \times \left[d-d_r\right]^{-d_r} \times \left[\left(\begin{array}{c} d \\ d_r \end{array} \right)\right]^{-1} \times [d+1]^{-1},
\end{eqnarray}
where $\left(\begin{array}{c} a \\ b \end{array}\right)$ designs combination of $a$ things taken $b$ at a time without repetition and where all probabilities $\mathbb{P}_U(\cdot)$ denote uniform distribution on the related space at hand. $\mathbb{P}_H(\boldsymbol{S})$ gives decreasing probabilities to more complex models, provided that the following new hypothesis is verified:

@@

This algorithm alternates two steps: the definition of a neighbourhood $\mathcal{V}(\boldsymbol{S})$ around the current structure $\boldsymbol{S}$ distribution and then the generation of a new structure $\boldsymbol{S}^+$ belonging to this neighbourhood according to its posterior probability.

@@

We define a {\it global} neighbourhood space $\mathcal{V}(\boldsymbol{S})$ of $\boldsymbol{S}$ composed by the following four {\it specific} neighbourhood spaces $\mathcal{V}(\boldsymbol{S})=\mathcal{V}_{r+}(\boldsymbol{S})\cup\mathcal{V}_{r-}(\boldsymbol{S})\cup\mathcal{V}_{p+}(\boldsymbol{S})\cup\mathcal{V}_{p-}(\boldsymbol{S})$ described below:
	\begin{itemize}
	\item {\bf Adding a sub-regression}: a new sub-regression with only one predictor covariate is added to $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{r+}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, (\tilde{\boldsymbol{J}_r},\tilde{\boldsymbol{J}_p})^{1,\ldots,d_r}=(\boldsymbol{J}_r,\boldsymbol{J}_p), \tilde{J}_r^{d_r+1}\in J_f, \tilde{\boldsymbol{J}}_p^{d_r+1}=\{j\}, j\in J_f \right\}.
\end{equation}	
		\item {\bf Removing a sub-regression}: a sub-regression with only one predictor covariate is removed from $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{r-}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, (\tilde{\boldsymbol{J}_r},\tilde{\boldsymbol{J}_p})=(\boldsymbol{J}_r,\boldsymbol{J}_p)^{\{1,\ldots,d_r\}\backslash j}, j\in\{1,\ldots,d_r\} \right\}.
\end{equation}	
	\item {\bf Adding a predictor covariate}: a predictor covariate is added to one sub-regression of $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{p+}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r, \tilde{\boldsymbol{J}}_p^{j}=\tilde{\boldsymbol{J}}_p^{j} \cup \{h\}, j\in \{1,\ldots,d_r\}, h\in J_f \right\}.
\end{equation}
	\item {\bf Removing a predictor covariate}: a predictor covariate is removed from one sub-regression of $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{p-}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r, \tilde{\boldsymbol{J}}_p^{j}=\tilde{\boldsymbol{J}}_p^{j} \backslash \{h\}, j\in \{1,\ldots,d_r\}, h\in J_f \right\}.
\end{equation}	
	\end{itemize}

@@

We then generate $\boldsymbol{S}^+$ from the following transition probability defined in  $\mathcal{V}(\boldsymbol{S})$
	\begin{equation}
	\boldsymbol{S}^+ \sim \frac{\exp{(-\mbox{{\sc bic}}_*(\cdot))}}{\sum_{\tilde{\boldsymbol{S}}\in \mathcal{V}(\boldsymbol{S})}\exp{(-\mbox{{\sc bic}}_*(\tilde{\boldsymbol{S}}))}}.
	\end{equation}

>>>>

However, $\mathcal{S}$ being combinatorial, $|\mathcal{S}|$ is huge. It has two cumulated consequences: First, the {\it exact} probability $\mathbb{P}(\boldsymbol{S}|\boldsymbol{X})$ may be of the same order of magnitude for a large number of candidates  $\boldsymbol{S}$, including the best one; Second, the {\sc bic} {\it approximations} of this quantity may introduce additional confusion to wisely distinguish between model probabilities. In order to limit this problem, we propose to introduce some information in $\mathbb{P}(\boldsymbol{S})$ promoting simple models through the following {\it hierarchical} uniform distribution denoted by $\mathbb{P}_H(\boldsymbol{S})$:
\begin{eqnarray}
\mathbb{P}_H(\boldsymbol{S}) & = & \mathbb{P}_H(\boldsymbol{J}_r,\boldsymbol{J}_p) \\
 & = & \mathbb{P}_H(\boldsymbol{J}_r,\boldsymbol{J}_p,d_r,\boldsymbol{d}_p) \\
 & = & \mathbb{P}_U(\boldsymbol{J}_p|\boldsymbol{d}_p,\boldsymbol{J}_r,d_r) \times \mathbb{P}_U(\boldsymbol{d}_p|\boldsymbol{J}_r,d_r) \times \mathbb{P}_U(\boldsymbol{J}_r|d_r)\times \mathbb{P}_U(d_r) \\
 & = & \left[\prod_{j=1}^{d_r} \left(\begin{array}{c} d-d_r \\ d_p^j \end{array}\right) \right]^{-1} \times \left[d-d_r\right]^{-d_r} \times \left[\left(\begin{array}{c} d \\ d_r \end{array} \right)\right]^{-1} \times [d+1]^{-1},
\end{eqnarray}
where $\left(\begin{array}{c} a \\ b \end{array}\right)$ designs the number of b-element subsets of an a-element set and where all probabilities $\mathbb{P}_U(\cdot)$ denote uniform distribution on the related space at hand. $\mathbb{P}_H(\boldsymbol{S})$ gives decreasing probabilities to more complex models, provided that the following new hypothesis is verified:

@@

This algorithm alternates two steps: the definition of a neighbourhood $\mathcal{V}(\boldsymbol{S})$ around the current structure $\boldsymbol{S}$ and then the generation of a new structure $\boldsymbol{S}^+$ belonging to this neighbourhood according to its posterior probability.

@@

We define a {\it global} neighbourhood space $\mathcal{V}(\boldsymbol{S})$ of $\boldsymbol{S}$ composed by the following four {\it specific} neighbourhood spaces $\mathcal{V}(\boldsymbol{S})=\mathcal{V}_{r+}(\boldsymbol{S})\cup\mathcal{V}_{r-}(\boldsymbol{S})\cup\mathcal{V}_{p+}(\boldsymbol{S})\cup\mathcal{V}_{p-}(\boldsymbol{S})$ described below:
	\begin{itemize}
	\item {\bf Adding a sub-regression}: a new sub-regression with only one predictor covariate is added to $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{r+}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, (\tilde{\boldsymbol{J}_r},\tilde{\boldsymbol{J}_p})^{1,\ldots,d_r}=(\boldsymbol{J}_r,\boldsymbol{J}_p), \tilde{J}_r^{d_r+1}\in J_f, \tilde{J}_p^{d_r+1}=\{j\}, j\in J_f \right\}.
\end{equation}	
		\item {\bf Removing a sub-regression}: a sub-regression with only one predictor covariate is removed from $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{r-}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, (\tilde{\boldsymbol{J}_r},\tilde{\boldsymbol{J}_p})=(\boldsymbol{J}_r,\boldsymbol{J}_p)^{\{1,\ldots,d_r\}\backslash j}, j\in\{1,\ldots,d_r\} \right\}.
\end{equation}	
	\item {\bf Adding a predictor covariate}: a predictor covariate is added to one sub-regression of $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{p+}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r, \tilde{J}_p^{j}=\tilde{J}_p^{j} \cup \{h\}, j\in \{1,\ldots,d_r\}, h\in J_f \right\}.
\end{equation}
	\item {\bf Removing a predictor covariate}: a predictor covariate is removed from one sub-regression of $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{p-}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r, \tilde{J}_p^{j}=\tilde{J}_p^{j} \backslash \{h\}, j\in \{1,\ldots,d_r\}, h\in J_f \right\}.
\end{equation}	
	\end{itemize}

@@

We then generate $\boldsymbol{S}^+$ from the following transition probability defined in  $\mathcal{V}(\boldsymbol{S})$
	\begin{equation}
	\mathbb{P}(\boldsymbol{S}^+|\mathcal{V}(\boldsymbol{S})) \sim \frac{\exp{(-\mbox{{\sc bic}}_*(\cdot))}}{\sum_{\tilde{\boldsymbol{S}}\in \mathcal{V}(\boldsymbol{S})}\exp{(-\mbox{{\sc bic}}_*(\tilde{\boldsymbol{S}}))}}.
	\end{equation}
